---
title: "Historical SAF planting analysis"
author: "Hardy Griesbauer"
date: "10/02/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                       fig.height=8,fig.width=12)
library(tidyverse)

options(scipen = 9999)

# Load data
load("../data/blData.RData")

# Some basic formatting
X<-
  blDat %>% 
   dplyr::select(Region=REGION,Num_planted=NUMBER_PLANTED,Year=YEAR,everything()) %>% 
    mutate(Region2=factor(Region,labels=c("Cariboo","NorthEast","Omineca","Skeena")))
                                         
      

```

## Historical analysis of SAF planting in northern BC

### Annual SAF planted by year in four regions

```{r}
X %>% 
  filter(SPECIES=="BL") %>%  # filter for BL only
  group_by(Region2,Year) %>% 
  summarise(Num_planted=sum(Num_planted)) %>% # summarize number of Bl planted
  ggplot()+
  aes(x=Year,y=Num_planted)+
  geom_bar(stat="Identity")+
  facet_wrap(~Region2)+
  theme(axis.text.x = element_text(angle = 270,vjust=0.3))+
  scale_x_continuous(breaks=seq(1985, 2020, 5),limits=c(1985, 2020))+
  ylab("Number of seedlings planted")



```

I kept the y-axis scaled the same for all regions to allow for better comparisons.

Figure shows that number of SAF planted in Skeena far exceeds number planted in other three regions.

Seems that SAF planting in all regions peaked between 1995 and 2000, and then dropped substantially around 2002.

We can zoom in on Omineca and Cariboo regions to see their numbers in more detail.

### Cariboo region
```{r}
X %>% 
  filter(SPECIES=="BL") %>%  # filter for BL only
  filter(Region2=="Cariboo") %>%  # filter for Cariboo region
  group_by(Region2,Year) %>% 
  summarise(Num_planted=sum(Num_planted)) %>% # summarize number of Bl planted
  ggplot()+
  aes(x=Year,y=Num_planted)+
  geom_bar(stat="Identity")+
  theme(axis.text.x = element_text(angle = 270,vjust=0.3))+
  scale_x_continuous(breaks=seq(1985, 2020, 5),limits=c(1985, 2020))+
  ylab("Number of seedlings planted")+
  ggtitle("Annual number of SAF seedlings planted in Cariboo Region")



```


### Omineca Region
```{r}
X %>% 
  filter(SPECIES=="BL") %>%  # filter for BL only
  filter(Region2=="Omineca") %>%  # filter for Cariboo region
  group_by(Region2,Year) %>% 
  summarise(Num_planted=sum(Num_planted)) %>% # summarize number of Bl planted
  ggplot()+
  aes(x=Year,y=Num_planted)+
  geom_bar(stat="Identity")+
  theme(axis.text.x = element_text(angle = 270,vjust=0.3))+
  scale_x_continuous(breaks=seq(1985, 2020, 5),limits=c(1985, 2020))+
  ylab("Number of seedlings planted")+
  ggtitle("Annual number of SAF seedlings planted in Omineca Region")
```

### Combined BA and BL planting in Skeena
Below is a figure showing annual planting in Skeena for both fir species (subalpine and amabalis).

```{r}
X %>% 
  filter(Region2=="Skeena") %>%  # filter for Cariboo region
  group_by(Region2,Year,SPECIES) %>% 
  summarise(Num_planted=sum(Num_planted)) %>% # summarize number of Bl planted
  ggplot()+
  aes(x=Year,y=Num_planted,fill=SPECIES)+
  geom_bar(stat="Identity",position="stack")+
  theme(axis.text.x = element_text(angle = 270,vjust=0.3))+
  scale_x_continuous(breaks=seq(1985, 2020, 5),limits=c(1985, 2020))+
  ylab("Number of seedlings planted")+
  ggtitle("Annual number of fir (BL/BA) seedlings planted in Skeena Region")


```

***
## Where are we planting SAF?
### Annual SAF planting by BGC unit categories
To show where SAF is being planted from an biogeoclimatic perspective, I created the following categories:

1. Wetbelt (ICH and SBSw/v subzones)
2. Sub-boreal/Boreal (SBS m/d subzones, BWBS); and 
3. High-elevation (ESSF)
4. Coastal (CWH/MH)


```{r}
X %>% 
  filter(SPECIES=="BL") %>% # filter for SAF
  mutate(BGC.Cat="Sub-Boreal") %>% # create new category column
  mutate(BGC.Cat=replace(BGC.Cat,Zone=="ESSF","High-El.")) %>%  # Create High-Elevation category
   mutate(BGC.Cat=replace(BGC.Cat,Zone=="SBS" & Subzone=="wk"|Subzone=="vk","Wetbelt")) %>%  # Create Wetbelt category
  mutate(BGC.Cat=replace(BGC.Cat,Zone=="ICH","Wetbelt")) %>%
  mutate(BGC.Cat=replace(BGC.Cat,Zone=="CWH"|Zone=="MH","Coastal")) %>% 
  drop_na(Zone) %>% # drop the one opening without BGC
  group_by(BGC.Cat,Year) %>% # group by BGC category
  summarise(Num_planted=sum(Num_planted)) %>%  # sum up planted numbers
  
# Produce plot
  ggplot()+
  aes(x=Year,y=Num_planted)+
  geom_bar(stat="Identity",position="stack")+
  theme(axis.text.x = element_text(angle = 270,vjust=0.3))+
  scale_x_continuous(breaks=seq(1985, 2020, 5),limits=c(1985, 2020))+
  facet_wrap(~BGC.Cat)+
  ylab("Number of seedlings planted")+
  ggtitle("Annual number of SAF seedlings planted by BGC category")


```

Most SAF seedlings in the 1990s were planted in wetbelt, but since ~2002 it look like equal amounts planted in sub-boreal, wetbelt and high-elevation units.

Just preliminary at this point.  I can send the data itself for you to create your own figures, or could change the figures to suit the presentation.  Whatever's easier.

More to come!

